---
title: "DC Differentiation sex classification"
output: rmarkdown::github_document
---

```{r message=FALSE}
## Inspired by cellXY : https://github.com/phipsonlab/cellXY
library(tidyverse)
library(speckle)
library(Seurat)
library(SeuratDisk)
library(glue)
library(ggplot2)
library(MASS)
library(pROC)
library(ggpubr)

```

To determine biological sex of the cells from gene expression, we wanted to use the most relevant cellular input for
a model. We decided that using DCs from ascites was the best data to create a model with. We used our ascites DCs and
used 5X cross-validation to see how accurate of a model we could create.

```{r test_model, warning = FALSE, message = FALSE}

assess.prediction=function(truth,predicted) {
  # check for missing values (we are going to # compute metrics on non-missing values only)
  predicted = predicted[ ! is.na(truth) ]
  truth = truth[ ! is.na(truth) ]
  truth = truth[ ! is.na(predicted) ]
  predicted = predicted[ ! is.na(predicted) ]
  # cat("Total cases that are not NA: ", length(truth),"\n",sep="")
  # overall accuracy of the test: how many cases # (both positive and # negative) we got right:
  # cat("Correct predictions (accuracy): ", sum(truth==predicted), "(",signif(sum(truth==predicted)*100/ length(truth),3),"%)\n",sep="")
  # how predictions align against known # training/testing outcomes: # TP/FP= true/false positives, # TN/FN=true/false negatives
  TP = sum(truth==1 & predicted==1)
  TN = sum(truth==0 & predicted==0)
  FP = sum(truth==0 & predicted==1)
  FN = sum(truth==1 & predicted==0)
  P = TP+FN  # total number of # positives in the truth data
  N = FP+TN  # total number of # negatives
  # cat("TPR (sensitivity)=TP/P: ", signif(100*TP/P,3),"%\n",sep="")
  # cat("TNR (specificity)=TN/N: ", signif(100*TN/N,3),"%\n",sep="")
  # cat("PPV (precision)=TP/(TP+FP): ", signif(100*TP/(TP+FP),3),"%\n",sep="")
  # cat("FDR (false discovery)=1-PPV: ", signif(100*FP/(TP+FP),3),"%\n",sep="")
  # cat("FPR =FP/N=1-TNR: ", signif(100*FP/N,3),"%\n",sep="")
  roc_object <- roc(truth, predicted)
  auroc <- auc(roc_object)

  # Add accuracy
  acc <- sum(truth == predicted) / length(truth) * 100

  return(c("sensitivity" = signif(100*TP/P,3),
           "specificity" = signif(100*TN/N,3),
           "auc" = auroc,
           "accuracy" = acc))
}
plot_auc_stats <- function(stat_df){
  plot_list <- list()
  for (v in c("sensitivity", "specificity", "accuracy")){
    median_val <- round(median(stat_df[[v]]), 2)
    p <- ggplot(stat_df, aes_string(y = v)) +
      geom_boxplot() +
      ggtitle(glue("{v}\n median={median_val}")) +
      theme_classic(base_size = 20) +
      theme(axis.text.x = element_blank(),
            axis.ticks.x = element_blank())
    plot_list <- c(plot_list, list(p))
  }
  return(plot_list)
}

# X chromosome genes that avoid inactivation
x_genes <- c("ARHGAP4","STS","ARSD", "ARSL", "AVPR2", "BRS3", "S100G", "CHM",
             "CLCN4", "DDX3X","EIF1AX","EIF2S3", "GPM6B", "GRPR", "HCFC1",
             "L1CAM", "MAOA", "MYCLP1", "NAP1L3", "GPR143", "CDK16", "PLXNB3",
             "PRKX", "RBBP7", "RENBP", "RPS4X", "TRAPPC2", "SH3BGRL", "TBL1X",
             "UBA1", "KDM6A", "XG", "XIST", "ZFX", "PUDP", "PNPLA4", "USP9X",
             "KDM5C", "SMC1A", "NAA10", "OFD1", "IKBKG", "PIR", "INE2", "INE1",
             "AP1S2", "GYG2", "MED14", "RAB9A", "ITM2A", "MORF4L2", "CA5B",
             "SRPX2", "GEMIN8", "CTPS2", "CLTRN", "NLGN4X", "DUSP21", "ALG13",
             "SYAP1", "SYTL4", "FUNDC1", "GAB3", "RIBC1", "FAM9C","CA5BP1")
y_genes <-  c("AMELY", "DAZ1", "PRKY", "RBMY1A1", "RBMY1HP", "RPS4Y1", "SRY",
            "TSPY1", "UTY", "ZFY","KDM5D", "USP9Y", "DDX3Y", "PRY", "XKRY",
            "BPY2", "VCY", "CDY1", "EIF1AY", "TMSB4Y","CDY2A", "NLGN4Y",
            "PCDH11Y", "HSFY1", "TGIF2LY", "TBL1Y", "RPS4Y2", "HSFY2",
            "CDY2B", "TXLNGY","CDY1B", "DAZ3", "DAZ2", "DAZ4")
all_genes <- c(x_genes, y_genes) %>% unique()

### Build model with ascites DCs ###
# Get the ascites DC data
data_path <- "/projects/home/nealpsmith/projects/ascites/dc_diff_3/sex_classification/data/ascites_dc_R8_300mg_20pm_harm_channel_1_1_complete_with_pb"
# Convert(paste(data_path, ".h5ad", sep = ""), des = "h5seurat", )
all_data <- LoadH5Seurat(paste(data_path, ".h5seurat", sep = ""), meta.data = FALSE, misc = FALSE)

# Read in obs
obs <- read.csv("/projects/home/nealpsmith/projects/ascites/dc_diff_3/sex_classification/data/ascites_dc_R8_300mg_20pm_harm_channel_1_1_complete_with_pb_obs.csv", row.names = 1)

norm_res <- all_data@assays$RNA@data
keep_genes <- all_genes[all_genes %in% rownames(norm_res)]

# Make a dataframe for logistic regression
regression_df <- norm_res[keep_genes,] %>%
  as.matrix() %>%
  t() %>%
  as.data.frame() %>%
  rownames_to_column("cell_id") %>%
  dplyr::left_join(obs %>% dplyr::select("sex", "tissue_type") %>% rownames_to_column("cell_id"), by = "cell_id") %>%
  dplyr::filter(tissue_type == "ascites") %>%
  column_to_rownames("cell_id") %>% dplyr::filter(!is.na(sex)) %>%
  dplyr::select(-tissue_type)

# Remve things not well detected
keep_col <- colSums(regression_df[,-ncol(regression_df)]) > 1000
keep_col <- names(keep_col[keep_col == TRUE])
regression_df <- regression_df %>% dplyr::select(c(keep_col, "sex"))


regression_df$sex <- ifelse(regression_df$sex == "M", 1, 0)

# 5X CV for looking at accuracy
regression_df$cv_group <- sample(1:nrow(regression_df) %% 5)

eval_df <- data.frame()
for (i in unique(regression_df$cv_group)){
  train_dat <- regression_df %>%
    dplyr::filter(cv_group != i) %>%
    dplyr::select(-cv_group)
  test_dat <- regression_df %>%
    dplyr::filter(cv_group == i) %>%
    dplyr::select(-sex) %>%
    dplyr::select(-cv_group)
  # test_dat <- scale(test_dat) %>% as.data.frame()
  # train_dat[,-c(ncol(train_dat))] <- scale(train_dat[,-c(ncol(train_dat))]) %>% as.data.frame()
  logreg_model <- glm(sex~. , family = "binomial", data = train_dat)

  preds <- predict(logreg_model, newdata = test_dat, type = "response") %>%
    round() %>%
    as.data.frame() %>%
    `colnames<-`(c("predicted_sex")) %>%
    rownames_to_column("cell_id") %>%
    dplyr::left_join(obs %>% dplyr::select("sex") %>% rownames_to_column("cell_id"), by = "cell_id")
  preds$sex <- ifelse(preds$sex == "M", 1, 0)
  eval <- assess.prediction(preds$sex, preds$predicted_sex)
  df <- as.data.frame(eval) %>% t() %>%
        as.data.frame() %>%
        mutate(iter = i)
  eval_df <- rbind(eval_df, df)

}
plot_list <- plot_auc_stats(eval_df)
ggarrange(plotlist = plot_list)

```

So our Logistic regression model was 90% accurate. We decided to go forward with this model and predict the sex of the
cells from our DC differentiation experiment using all ascites DCs as our training set

```{r create_predictions, message = FALSE, warning = FALSE}

### Okay we have ~90% accuracy, lets train on all DCs, apply to differentiation experiment ###
data_path <- "/projects/home/nealpsmith/projects/ascites/dc_diff_3/sex_classification/data/ascites_dc_R8_300mg_20pm_harm_channel_1_1_complete_with_pb"
# Convert(paste(data_path, ".h5ad", sep = ""), des = "h5seurat", )
ascites_dc_data <- LoadH5Seurat(paste(data_path, ".h5seurat", sep = ""), meta.data = FALSE, misc = FALSE)

# Read in obs
ascites_obs <- read.csv("/projects/home/nealpsmith/projects/ascites/dc_diff_3/sex_classification/data/ascites_dc_R8_300mg_20pm_harm_channel_1_1_complete_with_pb_obs.csv", row.names = 1)

ascites_norm_res <- ascites_dc_data@assays$RNA@data

# Now load the differentiation data
dc_diff_path <- "/projects/home/nealpsmith/projects/ascites/dc_diff_3/sex_classification/data/dc_differentiation_processed_data"
# Convert(paste(dc_diff_path, ".h5ad", sep = ""), des = "h5seurat")
dc_diff_data <- LoadH5Seurat(paste(dc_diff_path, ".h5seurat", sep = ""), meta.data = FALSE, misc = FALSE)

diff_norm_res <- dc_diff_data@assays$RNA@data

common_genes <- intersect(rownames(diff_norm_res), rownames(ascites_norm_res))

keep_genes <- all_genes[all_genes %in% common_genes]

train_df <- ascites_norm_res[keep_genes,] %>%
  as.matrix() %>%
  t() %>%
  as.data.frame() %>%
  rownames_to_column("cell_id") %>%
  dplyr::left_join(ascites_obs %>% dplyr::select("sex", "tissue_type") %>% rownames_to_column("cell_id"), by = "cell_id") %>%
  dplyr::filter(tissue_type == "ascites") %>%
  column_to_rownames("cell_id") %>% dplyr::filter(!is.na(sex)) %>%
  dplyr::select(-tissue_type)

train_df$sex <- ifelse(train_df$sex == "M", 1, 0)

test_df <- diff_norm_res[keep_genes,] %>%
  as.matrix() %>%
  t() %>%
  as.data.frame()

logreg_model <- glm(sex~. , family = "binomial", data = train_df)

preds <- predict(logreg_model, newdata = test_df, type = "response") %>%
    round() %>%
    as.data.frame() %>%
    `colnames<-`(c("predicted_sex"))
preds$predicted_sex <- ifelse(preds$predicted_sex == 1, "M", "F")
#
# write.csv(preds, "/projects/home/nealpsmith/projects/ascites/dc_diff_3/sex_classification/data/dc_diff_predicted_sex_for_cells_train_ascites_only.csv")


```